---
title: "PGA Golf DFS League Analytics"
author: "C.J. Tuskan"
date: last-modified
subtitle: "Testing"
format: html
editor: visual
df-print: kable
embed-resources: true
execute: 
  echo: true
toc: true
---

# Report Building

## Get DFS Leaderboard

## Get Race to 10 events (line plot to see trends)

## Get top 5 selected golfers (POPULAR LEAGUE PICKS)

## Get top 3 value golfers (CALC VALUE)

## Get bottom 3 value golfers

-   possibly a bubble plot for the above 2? Showing golfer selection volume and avg value?

## made cut percentage?

## Get wake-up call to users with most missed weeks

## User specific top pics and value picks and made cut percentage

# Predictions

## Current year / Current Event

```{r}
#current year schedule
df_sched <- fromJSON(paste('https://feeds.datagolf.com/get-schedule?tour=pga&file_format=json&key=', api_key, sep = ''))
df_sched <- data.frame(df_sched) %>% clean_names()

df_sched %>% mutate(schedule_start_date = as.Date(schedule_start_date)) 
```

```{r}
#upcoming round
df_field_updates <- fromJSON(paste('https://feeds.datagolf.com/field-updates?tour=pga&file_format=json&key=', api_key, sep = ''))
df_field_updates
```
