---
title: "PGA Course Analysis"
author: 'C.J. Tuskan'
date: last-modified
format:
 html:
   grid:
      sidebar-width: 1px
      body-width: 1200px
      margin-width: 300px
      gutter-width: 1.5rem
editor: visual
df-print: kable
embed-resources: true
execute: 
  echo: true
toc: true
---

# Ideas

-   why limiting to courses with 3+ events with last event in the past 2 years?
    -   Open it up to everything?
    -   Potentially bootstrap from everything

# Todo:

-   dig into the courses with singular fit
-   KNN similar courses

# Setup

```{r}
suppressWarnings(suppressPackageStartupMessages({
  library(tidyverse)
  library(ggplot2)
  library(janitor)
  library(jsonlite)
  library(lubridate)
  library(DBI)
  library(RPostgres)
  library(scales)
  library(RColorBrewer)
  library(ggimage)
  library(DT)
  library(mice)
  library(lme4)
  library(pROC)
  library(caret)
  library(randomForest)
  library(DataExplorer)
  library(car)
}))

#source db info
source('../../db_creds.R')
source('../../helper.R')

#color set
color_cjt <- c('#F1C40F', '#34495E', '#95A5A6')

# Establish a connection
conn <- dbConnect(
  drv = RPostgres::Postgres(),
  dbname = dbname,
  host = host,
  port = port,
  user = user,
  password = password
)

# Check the connection
if (!is.null(conn)) {
  print("Connection successful!")
} else {
  print("Connection failed!")
}
```

# Gather List of Courses

Filters:

-   3 or more events have been played at the course
-   the last event at the course was within the last year
-   the event at the course has a cut

```{r}
courses <- dbGetQuery(conn, "
with
    cte1 as (
        --individual golfer positions and dfs scoring grouped for each course
        select
            round.id_course,
            course.course_name,
            course.course_par,
            round.id_event,
            event.event_name,
            event.date,
            round.dg_id,
            dfs_total.fin_text,
            dfs_total.total_pts,
            CASE
                WHEN dfs_total.fin_text not in (' CUT', 'CUT') THEN 1
                ELSE 0
            END AS made_cut
        from
            round
            inner join course on course.id_course = round.id_course
            inner join event on event.id_event = round.id_event
            inner JOIN dfs_total on dfs_total.id_event = round.id_event
            and dfs_total.dg_id = round.dg_id
        where
            -- event.event_name like '%Masters%'
            -- event.calendar_year = 2025
            dfs_total.fin_text not in (
                'W/D',
                ' W/D',
                'WD',
                'DQ',
                'MDF',
                ' WD'
            )
        GROUP BY
            round.id_course,
            course.course_name,
            course.course_par,
            round.id_event,
            event.event_name,
            event.date,
            round.dg_id,
            dfs_total.fin_text,
            dfs_total.total_pts
        ORDER BY dfs_total.total_pts desc
    ),
    cte2 as (
        -- total golfer count and made cut count grouped by course and year
        select
            id_course,
            course_name,
            course_par,
            id_event,
            event_name,
            date,
            count(distinct dg_id) as count_total_golfers,
            sum(made_cut) as count_made_cut,
            round(avg(total_pts)::NUMERIC, 1) as avg_dfs_points
        from cte1
            -- where event_name like '%Masters%'
        GROUP BY
            id_course,
            course_name,
            course_par,
            id_event,
            event_name,
            date
        ORDER BY date
    ),
    cte3 as (
        --count of events grouped by course and filtering out courses without a cut
        select
            id_course,
            course_name,
            course_par,
            count(distinct id_event) as count_events,
            count(distinct event_name) as count_event_name,
            max(extract(year from date)) as last_event,
            min(extract(year from date)) as first_event
        from cte2
        where count_total_golfers != count_made_cut
        and date < '2025-04-06' -- PRIOR TO VALERO TEXAS OPEN
        group by
            id_course,
            course_name,
            course_par
        order by count_events desc
    )
select * from cte3 where last_event in (2024, 2025) and count_events >= 3")

courses
```

# Find Strokes Gained Attributes

The courses are looped if strokes gained data can be imputed, it is, otherwise the data from that calendar year is dropped

```{r}
log_missing_values <- data.frame()
course_effects <- data.frame()
rounds_all_orig <- data.frame()
impute_threshold <- 0.33

##FOR EACH ROUND AT COURSE WITHIN CALENDAR YEAR

for(i in 1:length(courses$id_course)){
  c <- courses[i,]
  print(paste('working on ', c$course_name, sep = ''))

  rounds <- dbGetQuery(conn, paste("
                        SELECT
                        	ROUND.*,
                        	EVENT.CALENDAR_YEAR
                        FROM
                        	ROUND
                        	INNER JOIN EVENT ON EVENT.ID_EVENT = ROUND.ID_EVENT
                        WHERE
                        	ROUND.ID_COURSE = '", c$id_course,"'", sep = ''))
  
  #if there are missing values, impute as long as missing data is not higher than threshold in any sg column of interest or total metadata
  round_check <- rounds %>% 
    group_by(calendar_year) %>% 
    summarise(total_true = sum(across(everything(), ~ sum(is.na(.)))),
              total_false = sum(across(everything(), ~ sum(!is.na(.)))),
              total_sg_ott_true = sum(across(sg_ott, ~sum(is.na(.)))),
              total_sg_ott_false = sum(across(sg_ott, ~sum(!is.na(.)))),
              total_sg_app_true = sum(across(sg_app, ~sum(is.na(.)))),
              total_sg_app_false = sum(across(sg_app, ~sum(!is.na(.)))),
              total_sg_putt_true = sum(across(sg_putt, ~sum(is.na(.)))),
              total_sg_putt_false = sum(across(sg_putt, ~sum(!is.na(.))))
              ) %>% 
    mutate(prop_missing_total = total_true/(total_true+total_false)) %>% 
    mutate(prop_missing_sg_ott = total_sg_ott_true/(total_sg_ott_true + total_sg_ott_false)) %>% 
    mutate(prop_missing_sg_app = total_sg_app_true/(total_sg_app_true + total_sg_app_false)) %>% 
    mutate(prop_missing_sg_putt = total_sg_putt_true/(total_sg_putt_true + total_sg_putt_false)) %>% 
    cbind(id_course = c$id_course, course_name = c$course_name) %>% 
    select(id_course, course_name, calendar_year, prop_missing_sg_ott, prop_missing_sg_app, prop_missing_sg_putt, prop_missing_total) %>% 
    mutate(missing_a_lot = case_when(prop_missing_sg_app > impute_threshold | 
                                     prop_missing_sg_ott > impute_threshold | 
                                     prop_missing_sg_putt > impute_threshold ~ 1,
                                     .default = 0))
  
  #log missing value summary
  log_missing_values <- log_missing_values %>% rbind(round_check)
  
  #filter out years with a lot of missing data
  round_check <- round_check %>% filter(missing_a_lot == 0)
  rounds <- rounds %>% filter(calendar_year %in% round_check$calendar_year)
  
  #if sg data from all rounds are missing, skip this course
  if(length(rounds$id_round)==0){
    next
  }
  
  ## IMPUTE MISSING VALUES IF NECESSARY
  
  #sum the missing values from all sg columns
  if((round_check %>% summarise(sg_missing = sum(prop_missing_sg_ott)+sum(prop_missing_sg_app)+sum(prop_missing_sg_putt)))$sg_missing != 0){
    # Perform multiple imputation
    imputed_data <- suppressWarnings(mice(rounds, m = 5, method = "pmm", maxit = 5, seed = 1842, printFlag = F))
    
    # View completed dataset
    rounds <- complete(imputed_data)
  }
  
  ## ADD RESPONSE VARIABLE (WHETHER THE GOLFER MADE THE CUT OR NOT)
  
  finish <- dbGetQuery(conn, paste("
                        SELECT
                        	DFS_TOTAL.ID_EVENT,
                        	DFS_TOTAL.DG_ID,
                        	DFS_TOTAL.FIN_TEXT,
                        	EVENT.CALENDAR_YEAR,
                        	EVENT.DATE
                        FROM
                        	DFS_TOTAL
                        	INNER JOIN EVENT ON EVENT.ID_EVENT = DFS_TOTAL.ID_EVENT
                        WHERE
                        	DFS_TOTAL.ID_EVENT IN (
                        		SELECT
                        			ID_EVENT
                        		FROM
                        			EVENT
                        		WHERE
                        			ID_EVENT IN ", get_sql_list(unique(rounds$id_event)),"
                        	)  
                        ", sep = ''))
  
  #add made_cut column
  finish_with_cut <- finish %>% 
    # filter(!fin_text %in% c('W/D', ' W/D', 'WD', 'DQ', 'MDF', ' WD')) %>% 
    mutate(made_cut = ifelse(!fin_text %in% c('CUT', ' CUT'), 1, 0))
  
  #add in whether made cut
  rounds <- rounds %>% 
    inner_join(finish_with_cut %>% 
                select(dg_id, calendar_year, date, id_event, fin_text, made_cut), by = c('calendar_year', 'dg_id', 'id_event'), relationship = 'many-to-one')
  
  #filter out withdraws, dq's, etc
  rounds <- rounds %>% 
    filter(!fin_text %in% c('W/D', ' W/D', 'WD', 'DQ', 'MDF', ' WD'))
  
    #add made_cut flag
  rounds <- rounds %>% 
    # mutate(made_cut = ifelse(!fin_text %in% c('CUT', ' CUT'), 1, 0)) %>% 
    select(-fin_text) %>%
    mutate(round = factor(round))
  
  #add to rounds_all df
  cat('   ', length(unique(rounds$id_event)), ' unique events\n')
  # for(e in 1:length(unique(rounds$id_event))){
  #   cat(unique(rounds$id_event)[e], '\n')
  # }
  rounds_all_orig <- rounds_all_orig %>% rbind(rounds)
  
  #filter to rounds 1 and 2 for course analysis
  rounds <- rounds %>%
    filter(round %in% c(1,2))
  
  ## SPLIT INTO TRAINING VALIDATION AND TEST SET
  
  # train <- rounds %>% filter(! calendar_year %in% c(2024, 2025))
  # val <- rounds %>% filter(calendar_year == 2024)
  # test <- rounds %>% filter(calendar_year == 2025)
  
  ## GENERATE NULL AND MIXED MODEL
  
  mod_null <- glmer(made_cut ~ 1 + (1|calendar_year:dg_id), family = binomial, data = rounds)
  mod <- glmer(made_cut ~ sg_putt + sg_ott + sg_app + (1|calendar_year:dg_id), family = binomial, data = rounds)
  s <- summary(mod)
  
  ## LOG THE course_effects
  
  course_effects <- course_effects %>% rbind(data.frame(id_course = c$id_course,
                                           course_name = c$course_name,
                                           aic_null = AIC(mod_null), 
                                           aic_mod = AIC(mod),
                                           logit_sg_putt = s$coefficients[2,1],
                                           logit_sg_ott = s$coefficients[3,1],
                                           logit_sg_app = s$coefficients[4,1],
                                           p_sg_putt = sg_putt <- s$coefficients[2,4],
                                           p_sg_ott = sg_ott <- s$coefficients[3,4],
                                           p_sg_app = sg_app <- s$coefficients[4,4],
                                           randef_var = (as.data.frame(VarCorr(mod)))$vcov))
}
```

```{r}
course_effects
```

# EDA

```{r}
plot_histogram(rounds_all_orig)
```

```{r}
plot_bar(rounds_all_orig)
```

```{r}
plot_correlation(rounds_all_orig, cor_args = list(use = "pairwise.complete.obs"))
```

```{r}
rounds_all_orig
```

```{r}
#ensure all fixed effects are significant and model is not Singular
p_thresh <- .05
course_effects_filtered <- course_effects %>% 
  filter(randef_var>10) %>% 
  filter(p_sg_putt<=p_thresh & p_sg_ott<=p_thresh & p_sg_app<=p_thresh)
course_effects_filtered
```

# Find Similar Courses via KNN

```{r}
course_effects_of_interest <- course_effects_filtered %>% filter(id_course == 'cf8c066a-2b81-4b7f-a661-d98e2774955a')

euch_dist <- sqrt(((course_effects_filtered$logit_sg_app - course_effects_of_interest$logit_sg_app)^2 +
                     (course_effects_filtered$logit_sg_putt - course_effects_of_interest$logit_sg_putt)^2 +
                     (course_effects_filtered$logit_sg_ott - course_effects_of_interest$logit_sg_ott)^2))

course_effects_filtered$similarity <- euch_dist

# get top similar courses
list_similar_courses <- (course_effects_filtered %>% arrange(similarity) %>% head(6) %>% tail(5))$id_course
list_similar_courses
```

```{r}
# set calendar for event of interest
# e.g. valero texas open in april 2022 -> want to utilize 1 year worth of data leading up to that event.  
#   - if grouped by calendar_year, this shifts potential input data for validation and test sets
#   - need to feature engineer the start/end of 'event of interest' ('eoi') year
#   - calendar_year ==  2022 for Valero, feature engineered column will assign 2022 for all events after Valero 2021

#grab all event of interest with date and calendar year
events_of_interest <- dbGetQuery(conn, "
SELECT DISTINCT
    ROUND.id_event,
    course.id_course,
    course.course_name,
    EVENT.event_name,
    EVENT.calendar_year,
    EVENT.date
FROM course
    INNER JOIN ROUND ON ROUND.id_course = course.id_course
    INNER JOIN EVENT ON EVENT.id_event = ROUND.id_event
WHERE
    course.id_course = 'cf8c066a-2b81-4b7f-a661-d98e2774955a'
ORDER BY calendar_year")
events_of_interest
```

```{r}
# #copy rounds dataframe
# rounds_all <- rounds_all_orig
# 
# #gather event of interest year align right
# #e.g. 2023 events up to and including eoi in 2023 will be assigned 2023
# rotated_year_right <- rounds_all %>%
#   group_by(id_event, calendar_year, date) %>%
#   summarize(golfer_count = n(), .groups = 'drop') %>%
#   mutate(eoi_year_right = cut(date,
#                         breaks = events_of_interest$date+1,
#                         labels = (events_of_interest$calendar_year) %>% tail(length(events_of_interest$calendar_year)-1))) %>%
#   mutate(eoi_year_right = case_when(is.na(eoi_year_right) & calendar_year == min(events_of_interest$calendar_year) ~ factor(min(events_of_interest$calendar_year)), .default = eoi_year_right))
# 
# #gather event of interest year align left
# #e.g. 2022 eoi up to (but not including) the 2023 eoi will be assigned 2023
# rotated_year_left <- rounds_all %>%
#   group_by(id_event, calendar_year, date) %>%
#   summarize(golfer_count = n(), .groups = 'drop') %>%
#   mutate(eoi_year_left = cut(date,
#                         breaks = events_of_interest$date-1,
#                         labels = (events_of_interest$calendar_year) %>% tail(length(events_of_interest$calendar_year)-1))) %>%
#   mutate(eoi_year_left = case_when(is.na(eoi_year_left) & calendar_year == min(events_of_interest$calendar_year) ~ factor(min(events_of_interest$calendar_year)), .default = eoi_year_left))
# 
# rounds_all <- rounds_all %>%
#   left_join(rotated_year_right %>% select(id_event, eoi_year_right), by = 'id_event', relationship = 'many-to-one') %>%
#   left_join(rotated_year_left %>% select(id_event, eoi_year_left), by = 'id_event', relationship = 'many-to-one') %>% 
#   filter(!is.na(eoi_year_left) & !is.na(eoi_year_right))
# 
# #UTILIZED FOR EOI ALIGNED LEFT - ORIGINALLY WAS GOING TO UTILIZE TO PREDICT NEXT EOI WITH FLAG THAT SPANNED THE ENTIRE DF WHICH OVERFITS
# #add column for event of interest - made next cut
# rounds_all <- rounds_all %>%
# 
#   #debug filters
#   # rounds_all %>%
#   # filter(dg_id %in% c(4163, 7960)) %>%
#   # filter(id_event %in% events_of_interest$id_event) %>%
#   # select(id_event, dg_id, round, calendar_year, date, made_cut, eoi_year_left, eoi_year_right) %>% arrange(dg_id, date) %>%
# 
#   #add flag for eoi and its year
#   mutate(eoi_prelim_year = eoi_year_left) %>%
#   mutate(is_eoi = ifelse(id_event %in% events_of_interest$id_event, 1, 0)) %>%
# 
#   #get the made_cut value for the grouped eoi year
#   group_by(dg_id, eoi_year_left) %>%
#   mutate(made_cut_eoi_left = first(made_cut[id_event %in% events_of_interest$id_event])) %>%
#   ungroup() %>%
#     
#   #get the next eoi made_cut value
#   arrange(dg_id, date) %>%
#   group_by(dg_id) %>%
#   mutate(made_cut_next_eoi = lead(made_cut_eoi_left)) %>%
#   ungroup() %>%
#   
#   #set to entire eoi_year grouping to the lead made_cut value
#   group_by(dg_id, eoi_year_left) %>%
#   mutate(made_cut_next_eoi = last(made_cut_next_eoi)) %>%
#   ungroup() %>%
# 
#   #order rounds prior to eoi (1 = most recently played prior to eoi)
#   group_by(dg_id, eoi_year_left) %>%
#   mutate(eoi_recent_round_order = dense_rank(desc(date))) %>%
#   ungroup() %>%
# 
#   #add weights (in reverse order to ensure most important weights are applied if overwriting)
#   #e.g. there is a similar course that is played 2 rounds before event of interest
#   mutate(weighting = case_when(eoi_recent_round_order %in% c(1,2,3) & !is.na(made_cut_next_eoi) ~ 2.0, #recent performance prior to event of interest
#                                id_course %in% list_similar_courses ~ 3.0, #performance at similar courses
#                                is_eoi == 1 ~ 4.0, #previous performance at course
#                                .default = 1.0)) %>% 
# 
#   #remove rows where the golfer did not participate in the event of interest during a given year
#   filter(!(is.na(made_cut_next_eoi))) %>%
# 
#   #cleanup
#   select(-eoi_year_left, -eoi_year_right, -made_cut_eoi_left) %>%
#   arrange(dg_id, eoi_prelim_year, date)
# 
# rounds_all %>% arrange(dg_id, date)
```

```{r}
#copy rounds dataframe
rounds_all <- rounds_all_orig

#THIS SECTION WAS FOR DF TRAINING/PREDICTING RESULTS WITH ROUND LEVEL DATA FOR EOI
#EOI ALIGN RIGHT APPROACH
#THIS DOESN'T SEEM TO WORK BECAUSE IT TRAINS THE MODEL TO PREDICT HEAVILY ON WHAT THE SG VALUES ARE WHICH IS NOT KNOWN PRIOR TO TOURNAMENT
#I INITIALLY TRIED FILLING IN AVG VALUES FOR SG THINKING TRAINING AND PRIOR EVENTS WOULD HELP, BUT IT OVERFITS
#rank most recent events

#gather event of interest year align right
#e.g. 2022/2023 events up to and including eoi in 2023 will be assigned 2023
rotated_year_right <- rounds_all %>%
  group_by(id_event, calendar_year, date) %>%
  summarize(golfer_count = n(), .groups = 'drop') %>%
  mutate(eoi_year_right = cut(date,
                        breaks = events_of_interest$date+1, # (set to -1 for rotated left)
                        labels = (events_of_interest$calendar_year) %>% tail(length(events_of_interest$calendar_year)-1))) %>%
  mutate(eoi_year_right = case_when(is.na(eoi_year_right) & calendar_year == min(events_of_interest$calendar_year) ~ factor(min(events_of_interest$calendar_year)), 
                                    is.na(eoi_year_right) & calendar_year == max(events_of_interest$calendar_year) ~ factor(max(events_of_interest$calendar_year) + 1),
                                    .default = eoi_year_right))

rounds_all <- rounds_all %>%
  left_join(rotated_year_right %>% select(id_event, eoi_year_right), by = 'id_event', relationship = 'many-to-one') %>%
  # left_join(rotated_year_left %>% select(id_event, eoi_year_left), by = 'id_event', relationship = 'many-to-one') %>%
  # filter(!is.na(eoi_year_left) & !is.na(eoi_year_right))

  #label each event of interest as 1, all others 0
  mutate(is_eoi = ifelse(id_event %in% events_of_interest$id_event, 1, 0)) %>%

  #add column of 'event of interest' date to be used below
  left_join(events_of_interest %>%
              select(calendar_year, date) %>%
              rename(eoi_date = date, 
                     eoi_year_right = calendar_year) %>%
              mutate(eoi_year_right = factor(eoi_year_right)), 
            by = 'eoi_year_right') %>%

  #group by golfer and arrange by date
  group_by(dg_id) %>%
  arrange(date) %>%

  #if the event date is prior to the 'event of interest' date (within 'event of interest' year), set true
  mutate(is_before_eoi = date < eoi_date) %>%

  #rank the order of events played prior to 'event of interest'.  e.g. 1 is most recent and will label the same for both rounds
  group_by(dg_id, eoi_year_right, is_before_eoi) %>%
  mutate(rank_before_eoi = ifelse(is_before_eoi,
                                 dense_rank(desc(date)),
                                 NA_integer_)) %>%

  #regroup
  ungroup() %>%
  group_by(dg_id, eoi_year_right) %>%

  #flag if the golfer played in the 'event of interest' that particular year
  #this is because 'recent round' weighting will not apply if the golfer didn't play that year
  mutate(played_in_eoi = ifelse(any(is_eoi == 1), 1, 0)) %>%
  arrange(dg_id,date) %>%
  ungroup() %>%

#weight certain scenarios
# rounds_all <- rounds_all %>%
  mutate(weighting = case_when(id_event %in% events_of_interest$id_event ~ 4.0, #previous performance at course
                               id_course %in% list_similar_courses ~ 3.0, #performance at similar courses
                               rank_before_eoi %in% c(1,2,3) & played_in_eoi == 1 ~ 2.0, #recent performance prior to event of interest
                               .default = 1.0)) %>%
  
  #standardize
  group_by(eoi_year_right, id_event, round) %>% 
  mutate_at(c("driving_dist"), ~(scale(.) %>% as.vector)) %>% 
  mutate_at(c("driving_acc"), ~(scale(.) %>% as.vector)) %>% 
  mutate_at(c("gir"), ~(scale(.) %>% as.vector)) %>% 
  mutate_at(c("scrambling"), ~(scale(.) %>% as.vector)) %>% 
  mutate_at(c("prox_rgh"), ~(scale(.) %>% as.vector)) %>% 
  mutate_at(c("prox_fw"), ~(scale(.) %>% as.vector)) %>% 
  ungroup() %>% 

  #clean up
  mutate(eoi_year = eoi_year_right) %>%
  select(-eoi_year_right, -rank_before_eoi) %>% 
  
  #ADDING THIS FOR QUICK TEST - MAY WANT TO IMPUTE
  na.omit()

# rounds_all %>% filter(id_event %in% events_of_interest$id_event) %>% filter(dg_id %in% c(4163, 5665))

# summary(is.na(rounds_all))

rounds_all
```

# train/val/test

```{r}
# rounds_all_agg_mean <- rounds_all %>%
#   
#   #agg to event
#   group_by(dg_id, id_event) %>%
#   reframe(eoi_recent_round_order = first(eoi_recent_round_order),
#            weighting = first(weighting),
#            eoi_prelim_year = first(eoi_prelim_year),
#            made_cut = first(made_cut),
#            made_cut_next_eoi = first(made_cut_next_eoi),
#            date = first(date),
#            sg_app = mean(sg_app),
#            sg_putt = mean(sg_putt),
#            sg_ott = mean(sg_ott)) %>%
#   arrange(dg_id, eoi_prelim_year, date) %>%
#   ungroup() 
#   
#   #agg to year
#   group_by(dg_id, eoi_prelim_year) %>%
#   reframe(made_cut_next_eoi = first(made_cut_next_eoi),
#           sg_app = sum(sg_app*weighting)/sum(weighting),
#           sg_putt = sum(sg_putt*weighting)/sum(weighting),
#           sg_ott = sum(sg_ott*weighting)/sum(weighting),
#           made_cut_prior = sum(made_cut*weighting)/sum(weighting)) %>% 
#   ungroup()



#split into test/validate/train
train <- rounds_all %>% filter(! eoi_year %in% c(2024, 2025))
  val <- rounds_all %>% filter(eoi_year == 2024)
  test <- rounds_all %>% filter(eoi_year == 2025)
  
# val_actual <- rounds_all_orig %>% filter(id_event %in% events_of_interest$id_event, calendar_year == 2023) %>% 
#   group_by(dg_id) %>% 
#   summarize(made_cut = first(made_cut))

# test_actual <- rounds_all_orig %>% filter(id_event %in% events_of_interest$id_event, calendar_year == 2025) %>% 
#   group_by(dg_id) %>% 
#   summarize(made_cut = first(made_cut))
```

```{r}
#fit null model
mod_sg_ott_null <- lmer(sg_ott ~ 1 + (1|dg_id) + (1|dg_id:eoi_year), data = train, weights = weighting)
summary(mod_sg_ott_null)
```

```{r}
#fit model
mod_sg_ott <- lmer(sg_ott ~ driving_dist + driving_acc + gir + sg_t2g + sg_total + (1|dg_id) + (1|dg_id:eoi_year), 
                   data = train, 
                   weights = weighting)
summary(mod_sg_ott)
```

```{r}
#validation of model
anova(mod_sg_ott, mod_sg_ott_null)
```

```{r}
vif_mod_sg_ott <- lm(sg_ott ~ driving_dist + driving_acc + gir + sg_t2g + sg_total, 
                   data = train, 
                   weights = weighting)

vif(vif_mod_sg_ott)
```

```{r}
plot(sample(vif_mod_sg_ott$fitted.values, 1000), sample(vif_mod_sg_ott$residuals, 1000))
```

```{r}
mod_sg_ott_all <- lm(sg_ott ~ ., 
                   # data = train %>% select(-id_round, -id_event, -id_course, -dg_id, -start_hole, -teetime, -made_cut, -date, -is_eoi, -eoi_date, -is_before_eoi, -rank_before_eoi, -played_in_eoi, -calendar_year, -eoi_year), 
                   data = train %>% select(driving_dist, driving_acc, gir, sg_t2g, sg_total, weighting, sg_ott), 
                   weights = weighting)

step(mod_sg_ott_all)
```

```{r}
# summary(mod_sg_ott)

val_eoi <- (events_of_interest %>% filter(calendar_year == 2024))
val_eoi_sg_ott_actual <- rounds_all_orig %>% 
  filter(id_event == val_eoi$id_event) %>% 
  select(dg_id, id_event, round, sg_ott, made_cut) %>% 
  arrange(dg_id) %>% 
  rename(sg_ott_actual = sg_ott) %>% 
  group_by(dg_id, id_event, made_cut) %>% 
  summarize(sg_ott_actual = mean(sg_ott_actual), .groups = 'drop') %>%  
  arrange((sg_ott_actual))

val_eoi_prior <- rounds_all %>% 
  filter(date<val_eoi$date) %>%
  filter(dg_id %in% val_eoi_sg_ott_actual$dg_id) %>% 
  filter(weighting %in% c(2,4))

# val_eoi_prior_means <- rounds_all %>% 
#   filter(date<val_eoi$date) %>%
#   filter(dg_id %in% val_eoi_sg_ott_actual$dg_id) %>% 
#   filter(weighting %in% c(1,2,3,4)) %>%
#   select(id_event, dg_id, round, id_event, gir, driving_acc, driving_dist, scrambling, prox_rgh, prox_fw, sg_ott, sg_app, sg_putt, sg_t2g, sg_total, weighting, made_cut, eoi_year, made_cut, date) %>% 
#   group_by(id_event, date, dg_id) %>% 
#   summarize(gir = mean(gir),
#             driving_acc = mean(driving_acc),
#             driving_dist = mean(driving_dist),
#             scrambling = mean(scrambling),
#             prox_rgh = mean(prox_rgh),
#             prox_fw = mean(prox_fw),
#             sg_ott = mean(sg_ott),
#             sg_app = mean(sg_app),
#             sg_putt = mean(sg_putt),
#             sg_t2g = mean(sg_t2g),
#             sg_total = mean(sg_total),
#             weighting = mean(weighting), .groups = 'drop') %>% 
#   arrange(dg_id, date)

val_eoi_extraction <- val_eoi_prior %>%
  mutate(weighting = case_when(weighting == 2 ~ 10,
                               weighting == 3 ~ 20,
                               weighting == 4 ~ 50,
                               .default = 1)) %>% 
  mutate(prediction_date = date('2024-04-07')) %>% 
  group_by(dg_id) %>%
  nest() %>%
  mutate(model = map(data, ~ lm(sg_ott ~ date, data = ., weights = weighting))) %>%
  mutate(summary = map(model, glance), 
         event_count = map_int(data, nrow),
         sg_ott_pred = map(model, ~ predict(.x, newdata = data.frame(date = date('2024-04-07')))),
         sg_ott_mean = map_dbl(data, ~ mean(.x$sg_ott, na.rm = TRUE)),
         sg_ott_median = map_dbl(data, ~median(.x$sg_ott, na.rm = TRUE))) %>%
  unnest(sg_ott_pred) %>% 
  select(dg_id, event_count, sg_ott_pred, sg_ott_mean, sg_ott_median) %>% 
  mutate(sg_ott_pred = round(sg_ott_pred, 3)) %>% 
  mutate(sg_ott_mean = round(sg_ott_mean, 3)) %>% 
  mutate(sg_ott_median = round(sg_ott_median, 3)) %>% 
  full_join(val_eoi_sg_ott_actual, by = 'dg_id') %>% 
  mutate(sg_ott_pred_adj = ifelse(event_count < 7 | abs(sg_ott_pred) > 5, sg_ott_mean, sg_ott_pred)) 


# val_eoi_extraction_mean <- val_eoi_prior_means %>%
#   mutate(prediction_date = date('2024-04-07')) %>% 
#   group_by(dg_id) %>%
#   nest() %>%
#   mutate(model = map(data, ~ lm(sg_ott ~ date, data = .))) %>%
#   mutate(summary = map(model, glance), 
#          event_count = map_int(data, nrow),
#          sg_ott_pred = map(model, ~ predict(.x, newdata = data.frame(date = date('2024-04-07')))),
#          sg_ott_mean = map_dbl(data, ~ mean(.x$sg_ott, na.rm = TRUE))) %>%
#   unnest(sg_ott_pred) %>% 
#   # filter(dg_id == 25908)
#   select(dg_id, event_count, sg_ott_pred, sg_ott_mean) %>% 
#   mutate(sg_ott_pred = round(sg_ott_pred, 3)) %>% 
#   mutate(sg_ott_mean = round(sg_ott_mean, 3)) %>% 
#   full_join(val_eoi_sg_ott_actual, by = 'dg_id') %>% 
#   mutate(sg_ott_pred_adj = ifelse(event_count < 10 | abs(sg_ott_pred) > 3, sg_ott_mean, sg_ott_pred)) 

# a <- lm(sg_ott ~ date, data = val_eoi_extraction %>% filter(dg_id == 25908) %>% select(sg_ott, date))
# predict(a, newdata = data.frame(date = date('2024-04-07')))

val_eoi_extraction[is.na(val_eoi_extraction)] <- 0

rmse_mean <- sqrt(mean((val_eoi_extraction$sg_ott_actual - val_eoi_extraction$sg_ott_mean)^2))
rmse_pred <- sqrt(mean((val_eoi_extraction$sg_ott_actual - val_eoi_extraction$sg_ott_pred_adj)^2))

cat('RMSE mean:  ', rmse_mean)
cat('RMSE pred:  ', rmse_pred)
```

```{r}
ggplot(val_eoi_prior %>% filter(dg_id == 10423), aes(x = date, y = sg_ott, col = factor(weighting)))+
  geom_point()+
  geom_smooth(method = "lm", formula = y ~ 1, se = F, color = "red")+
  geom_smooth(method = "lm", se = F, color = "blue", formula = y ~ x)
```

```{r}
val_eoi_extraction %>% filter(dg_id == 10423)
```

```{r}

#get averages events (and similar events) for the year prior to validation
rounds_all_apprx_val <- rounds_all %>% 
  filter(date<(events_of_interest %>% filter(calendar_year == 2024 ))$date) %>%
  # filter(date>=(events_of_interest %>% filter(calendar_year == 2023))$date 
  #        & date<(events_of_interest %>% filter(calendar_year == 2024))$date) %>% 
  filter(weighting %in% c(3,4)) %>%
  mutate(weighting = ifelse(weighting == 3, 1, 2)) %>%
  # filter(weighting %in% c(4)) %>%
  # mutate(weighting = ifelse(weighting == 4, 1, NA)) %>%
  # group_by(dg_id, id_event) %>% 
  select(id_event, dg_id, round, id_event, gir, driving_acc, driving_dist, scrambling, prox_rgh, prox_fw, sg_ott, sg_app, sg_putt, sg_t2g, sg_total, weighting, made_cut, eoi_year, made_cut) %>% 
  group_by(dg_id, round) %>% 
  summarize(gir = mean(gir),
            driving_acc = mean(driving_acc),
            driving_dist = mean(driving_dist),
            scrambling = mean(scrambling),
            prox_rgh = mean(prox_rgh),
            prox_fw = mean(prox_fw),
            sg_ott = mean(sg_ott),
            sg_app = mean(sg_app),
            sg_putt = mean(sg_putt),
            sg_t2g = mean(sg_t2g),
            sg_total = mean(sg_total), .groups = 'drop') %>% 
  mutate(eoi_year = 2024) %>% 
  ungroup()

preds <- predict(mod_sg_ott, newdata = rounds_all_apprx_val, allow.new.levels = T)



rounds_all_apprx_val <- rounds_all_apprx_val %>% 
  mutate(sg_ott_predicted = preds) %>% 
  filter(dg_id %in% unique((rounds_all_orig %>% filter(id_event == val_eoi$id_event))$dg_id)) %>% 
  full_join(val_sg_ott_actual, by = c('dg_id', 'round'), relationship = 'one-to-one')
  # mutate(sg_ott_actual)


# get_sql_list(unique((rounds_all_apprx_val %>% filter(is.na(sg_ott)))$dg_id))
# get_sql_list(list_similar_courses)
# rounds_all_orig %>% filter(dg_id == 12909) %>% filter(id_event %in% events_of_interest$id_event)
train
```

```{r}



```

```{r}
mod_1_null <- glmer(made_cut_next_eoi ~ 1 + (1|dg_id), 
               family = binomial, 
               data = train,
               weights = weighting)
summary(mod_1_null)
```

```{r}
mod_1 <- glmer(made_cut ~ sg_ott + sg_app + sg_putt + (1|dg_id) + (1|dg_id:eoi_prelim_year), 
               family = binomial, 
               data = train,
               weights = weighting)
summary(mod_1)
```

```{r}
pred_probs <- predict(mod_1, newdata = val, type = "response", allow.new.levels = T)

cut_thresh <- 0.53

# val_pred <- val %>%
#   mutate(pred_probs = pred_probs) %>%
#   mutate(made_cut_next_eoi_predicted = ifelse(pred_probs>=cut_thresh, 1,0))


val_pred_agg <- val_pred %>%
  mutate(pred_probs = pred_probs) %>%
  group_by(dg_id) %>%
  summarize(made_cut_next_eoi = made_cut_next_eoi[1],
            pred_probs_combined_avg = mean(pred_probs),
            made_cut_next_eoi_predicted = ifelse(pred_probs_combined_avg>=cut_thresh, 1, 0)) %>%
  ungroup()

# new_probs <- val_approx %>% group_by(dg_id) %>% summarize(pred_probs_combined_avg = mean(pred_probs), made_cut_predicted = ifelse(pred_probs_combined_avg>=cut_thresh, 1, 0))
# 
# # val_approx <- val_approx %>% left_join(new_probs, by = 'dg_id', relationship = 'many-to-one')
# val_actual <- val %>% 
#   filter(is_eoi == 1) %>% 
#   group_by(dg_id) %>% 
#   summarize(made_cut = first(made_cut))
# 
# val_approx_agg <- val_approx_agg %>% left_join(val_actual, by = 'dg_id', relationship = 'many-to-one')
# 
# val_approx <- val_approx %>% left_join(val_approx_agg %>% select(dg_id, made_cut_predicted) %>% rename(made_cut = made_cut_predicted), by = 'dg_id', relationship = 'many-to-one')
  

auc_eoi <- roc(val_pred_agg$made_cut_next_eoi, val_pred_agg$pred_probs_combined_avg, levels = c(1,0), plot = T, direction = '>')$auc
```

# old

```{r}
#get averages of all prior events and similar events
rounds_all_apprx_val <- rounds_all %>% 
  # filter(date<(events_of_interest %>% filter(calendar_year == 2024 ))$date) %>% 
  filter(date< (val %>% filter(is_eoi == 1) %>% first())$date[1]) %>% 
  filter(weighting %in% c(3,4)) %>%
  mutate(weighting = ifelse(weighting == 3, 1, 2)) %>%
  # filter(weighting %in% c(4)) %>%
  # mutate(weighting = ifelse(weighting == 4, 1, NA)) %>%
  group_by(dg_id, id_event) %>% 
  select(dg_id, round, id_event, sg_ott, sg_app, sg_putt, weighting, made_cut, eoi_year, made_cut)

mod_1_approx <- glmer(made_cut ~ sg_ott + sg_app + sg_putt + (1|dg_id) + (1|dg_id:eoi_year), family = binomial, weights = weighting, data = rounds_all_apprx_val)
# mod_1_approx <- glmer(made_cut ~ sg_ott + sg_app + sg_putt + (1|dg_id), family = binomial, weights = weighting, data = rounds_all_apprx_val)
summary(mod_1_approx)
```

```{r}
#LEFT OFF HERE, NEED TO PREDICT MADE_CUT USING THE MODEL ABOVE WHEN FED AVG SG'S FROM THE TRAINING SET ABOVE, THEN TEST ACCURACY
#THIS IS ANOTHER POTENTIAL IMPROVEMENT TO THE ORIGINAL MODEL I BUILT BC IT INCLUDES SIMILAR COURSES IN THE MODEL
val_eoi <- (val %>% filter(is_eoi ==1) %>% first())$id_event[1]
val_golfers <- unique((val %>% filter(id_event == val_eoi))$dg_id)

#create val set sg averages
val_approx <- rounds_all_apprx_val %>% 
  group_by(dg_id, round) %>% 
  summarize(sg_ott = mean(sg_ott), 
            sg_app = mean(sg_app),
            sg_putt = mean(sg_putt),
            # made_cut = first(made_cut),
            .groups = 'drop') %>% 
  filter(dg_id %in% val_golfers) %>% 
  mutate(eoi_year = factor(2024)) %>% 
  mutate(id_event = val_eoi)

#NEED TO ACCOUNT FOR IF GOLFER DOESN'T EXIST
#ALSO COUNT OF GOLFERS IS REALLY LOW -- REVIEW THIS, THERE'S POTENTIAL!

pred_probs <- predict(mod_1_approx, newdata = val_approx, type = "response", allow.new.levels = T)

cut_thresh <- 0.53
val_approx_agg <- val_approx %>% 
  mutate(pred_probs = pred_probs) %>% 
  group_by(dg_id) %>% 
  summarize(pred_probs_combined_avg = mean(pred_probs), 
            made_cut_predicted = ifelse(pred_probs_combined_avg>=cut_thresh, 1, 0)) %>% 
  ungroup()

# new_probs <- val_approx %>% group_by(dg_id) %>% summarize(pred_probs_combined_avg = mean(pred_probs), made_cut_predicted = ifelse(pred_probs_combined_avg>=cut_thresh, 1, 0))

# val_approx <- val_approx %>% left_join(new_probs, by = 'dg_id', relationship = 'many-to-one')
val_actual <- val %>% 
  filter(is_eoi == 1) %>% 
  group_by(dg_id) %>% 
  summarize(made_cut = first(made_cut))

val_approx_agg <- val_approx_agg %>% left_join(val_actual, by = 'dg_id', relationship = 'many-to-one')

val_approx <- val_approx %>% left_join(val_approx_agg %>% select(dg_id, made_cut_predicted) %>% rename(made_cut = made_cut_predicted), by = 'dg_id', relationship = 'many-to-one')
  

auc_eoi <- roc(val_approx_agg$made_cut, val_approx_agg$pred_probs_combined_avg, levels = c(1,0), plot = T, direction = '>')$auc
```

```{r}
cm <- confusionMatrix(factor(val_approx_agg$made_cut_predicted), factor(val_approx_agg$made_cut), mode = "everything", positive = '1')

# Print results
print(cm)
```

```{r}
mod_null_1 <- glmer(made_cut ~ 1 + (1|dg_id) + (1|dg_id:eoi_year),
                    family = binomial,
                    data = train,
                    weights = weighting)
summary(mod_null_1)
```

```{r}
mod_1 <- glmer(made_cut ~ sg_putt + sg_ott + sg_app + (1|dg_id) + (1|dg_id:eoi_year),
               family = binomial,
               data = train,
               weights = weighting)
summary(mod_1)
```

```{r}
#incorporate approx eoi values into val set

val_eoi_snip <- val %>% filter(id_event == val_eoi)

val_new <- val %>% filter(id_event != val_eoi)

val_eoi_snip <- val_eoi_snip %>% 
  #remove approximated parameters
  select(-sg_ott, -sg_app, -sg_putt, -made_cut) %>% 
  
  #refill them with the approximates
  left_join(val_approx, by = c('id_event', 'eoi_year', 'dg_id', 'round'), relationship = 'one-to-one')

#check names with NA
#appear to be lesser known golfers or rookies in validation year
# dbGetQuery(conn, paste("select player_name from player where dg_id in ", get_sql_list(unique((val_eoi_snip %>% filter(is.na(made_cut)))$dg_id)), sep = ''))
# summary(is.na(val_eoi_snip))

#fill golfers with no prior history - set to 0 for no strokes gained and not making the cut
val_eoi_snip[is.na(val_eoi_snip)] <- 0

# val_eoi_snip$made_cut <- 0

#place the snippet back into the new val set
val_new <- rbind(val_new, val_eoi_snip)

val_new
```

```{r}
#VAL
pred_probs <- predict(mod_1, newdata = val, type = "response", allow.new.levels = T)

cut_thresh <- 0.53
# pred_probs_all <- val_new %>%
#   # select(id_event, dg_id, round, eoi_prelim_year, made_cut_next_eoi) %>%
#   select(id_event, dg_id, eoi_year, made_cut) %>%
#   mutate(pred_probs = pred_probs)
#   # mutate(made_cut_predicted = ifelse(pred_probs >=0.5, 1, 0)) 
# 
# pred_probs_all_agg <- pred_probs_all %>%
#   group_by(eoi_year, id_event, dg_id) %>%
#   summarize(made_cut = first(made_cut), pred_probs_avg = mean(pred_probs), pred_probs_max = max(pred_probs), .groups = 'drop') %>%
#   mutate(made_cut_predicted_for_avg = ifelse(pred_probs_avg>=cut_thresh, 1, 0),
#          made_cut_predicted_for_max = ifelse(pred_probs_max>=cut_thresh, 1, 0)) %>% 
#   mutate(made_cut_approx = made_cut) %>% 
#   select(-made_cut) %>% 
#   left_join(val_actual, by = 'dg_id', relationship = 'one-to-one')

pred_probs_eoi <- val %>%
  # select(id_event, dg_id, round, eoi_prelim_year, made_cut_next_eoi) %>%
  select(id_event, dg_id, eoi_year, made_cut) %>%
  mutate(pred_probs = pred_probs) %>%
  # mutate(made_cut_predicted = ifelse(pred_probs >=0.5, 1, 0)) %>%
  filter(id_event %in% events_of_interest$id_event)

pred_probs_eoi_agg <- pred_probs_eoi %>%
  group_by(eoi_year, id_event, dg_id) %>%
  summarize(made_cut = first(made_cut), pred_probs_avg = mean(pred_probs), pred_probs_max = max(pred_probs), .groups = 'drop') %>%
  mutate(made_cut_predicted_for_avg = ifelse(pred_probs_avg>=cut_thresh, 1, 0),
         made_cut_predicted_for_max = ifelse(pred_probs_max>=cut_thresh, 1, 0)) %>% 
  mutate(made_cut_actual = made_cut) %>% 
  select(-made_cut)
  # left_join(val_actual, by = 'dg_id', relationship = 'one-to-one')



#all
# auc <- roc(pred_probs_all_agg$made_cut, pred_probs_all_agg$pred_probs_avg, levels = c(1,0), plot = T, direction = '>')$auc

#eoi
# auc_eoi <- roc(pred_probs_eoi$made_cut, pred_probs_eoi$pred_probs, levels = c(1,0), plot = T, direction = '>')$auc
# auc_eoi <- roc(pred_probs_eoi_agg$made_cut, pred_probs_eoi_agg$pred_probs_avg, levels = c(1,0), plot = T, direction = '>')$auc
auc_eoi <- roc(pred_probs_eoi_agg$made_cut_actual, pred_probs_eoi_agg$pred_probs_avg, levels = c(1,0), plot = T, direction = '>')$auc

auc_eoi <- roc(val$made_cut, pred_probs, levels = c(1,0), plot = T, direction = '>')$auc
```

```{r}
cm <- confusionMatrix(factor(pred_probs_eoi_agg$made_cut_predicted_for_avg), factor(pred_probs_eoi_agg$made_cut), mode = "everything", positive = '1')

# Print results
print(cm)
```

```{r}
cat('percentage of golfers actually making the cut in the validation set:  ', mean(as.numeric(pred_probs_eoi$made_cut)), '\n')
cat('predicted percentage of golfers making the cut for validation input:  ', mean(pred_probs_eoi_agg$made_cut_predicted_for_avg), '\n')
# cat('ratio of golfers making the cut in the training data:  ', mean(as.numeric(train_1_avg$made_cut)), '\n')
cat('count of golfers in validation set (2024 tournament):  ', length(pred_probs_eoi_agg$dg_id), '\n')
cat('actual number of golfers to make the cut in the validation set:  ', sum(as.numeric(pred_probs_eoi_agg$made_cut)), '\n')
# cat('predicted number of golfers to make the cut in the validation set:  ', round(mean(df_preds$prediciton)*length(val$dg_id), 0))
cat('predicted number of golfers to make the cut in the validation set:  ', sum(as.numeric(pred_probs_eoi_agg$made_cut_predicted_for_avg)), '\n')
```

```{r}
#all
# df_preds <- as.data.frame(pred_probs) %>% 
#   mutate(prediciton = ifelse(pred_probs >= 0.50, 1, 0))
# 
# # Create confusion matrix
# cm <- confusionMatrix(factor(val$made_cut), factor(df_preds$prediciton), mode = "everything", positive = '1')
# 
# # Print results
# print(cm)

#eoi
# df_preds_eoi <- as.data.frame(pred_probs_eoi) %>% 
#   mutate(prediciton = ifelse(pred_probs >= 0.50, 1, 0))

# Create confusion matrix
# cm <- confusionMatrix(factor(pred_probs_eoi_agg$made_cut), factor(pred_probs_eoi_agg$made_cut_predicted_for_max), mode = "everything", positive = '1')
cm <- confusionMatrix(factor(pred_probs_eoi$made_cut_next_eoi), factor(pred_probs_eoi$made_cut_predicted), mode = "everything", positive = '1')

# Print results
print(cm)
```

```{r}
cat('percentage of golfers actually making the cut in the validation set:  ', mean(as.numeric(pred_probs_eoi$made_cut)), '\n')
cat('predicted percentage of golfers making the cut for validation input:  ', mean(pred_probs_eoi_agg$made_cut_predicted_for_avg), '\n')
# cat('ratio of golfers making the cut in the training data:  ', mean(as.numeric(train_1_avg$made_cut)), '\n')
cat('count of golfers in validation set (2024 tournament):  ', length(pred_probs_eoi_agg$dg_id), '\n')
cat('actual number of golfers to make the cut in the validation set:  ', sum(as.numeric(pred_probs_eoi_agg$made_cut)), '\n')
# cat('predicted number of golfers to make the cut in the validation set:  ', round(mean(df_preds$prediciton)*length(val$dg_id), 0))
cat('predicted number of golfers to make the cut in the validation set:  ', sum(as.numeric(pred_probs_eoi_agg$made_cut_predicted_for_avg)), '\n')
```

```{r}
#TEST

pred_probs <- predict(mod_1, newdata = test %>% select(sg_putt, sg_ott, sg_app, eoi_year, dg_id), type = "response", allow.new.levels = T)

pred_probs_eoi <- test %>% 
  select(id_event, dg_id, round, eoi_year, made_cut) %>% 
  mutate(pred_probs = pred_probs) %>% 
  mutate(made_cut_predicted = ifelse(pred_probs >=.5, 1, 0)) %>% 
  filter(id_event %in% events_of_interest$id_event)

cut_thresh <- 0.53
pred_probs_eoi_agg <- pred_probs_eoi %>% 
  group_by(eoi_year, id_event, dg_id) %>% 
  summarize(made_cut = first(made_cut), pred_probs_avg = mean(pred_probs), pred_probs_max = max(pred_probs), .groups = 'drop') %>% 
  mutate(made_cut_predicted_for_avg = ifelse(pred_probs_avg>=cut_thresh, 1, 0),
         made_cut_predicted_for_max = ifelse(pred_probs_max>=cut_thresh, 1, 0))

#all
# auc <- roc(test$made_cut, pred_probs, levels = c(1,0), plot = T, direction = '>')$auc

#eoi
# auc_eoi <- roc(pred_probs_eoi$made_cut, pred_probs_eoi$pred_probs, levels = c(1,0), plot = T, direction = '>')$auc
auc_eoi <- roc(pred_probs_eoi_agg$made_cut, pred_probs_eoi_agg$pred_probs_avg, levels = c(1,0), plot = T, direction = '>')$auc
```

```{r}
# Create confusion matrix
cm <- confusionMatrix(factor(pred_probs_eoi_agg$made_cut_predicted_for_avg), factor(pred_probs_eoi_agg$made_cut), mode = "everything", positive = '1')

# Print results
print(cm)

```

```{r}
cat('percentage of golfers actually making the cut in the test set:  ', mean(as.numeric(pred_probs_eoi$made_cut)), '\n')
cat('predicted percentage of golfers making the cut for test input:  ', mean(pred_probs_eoi_agg$made_cut_predicted_for_avg), '\n')
cat('count of golfers in test set (2024 tournament):  ', length(pred_probs_eoi_agg$dg_id), '\n')
cat('actual number of golfers to make the cut in the test set:  ', sum(as.numeric(pred_probs_eoi_agg$made_cut)), '\n')
cat('predicted number of golfers to make the cut in the test set:  ', sum(as.numeric(pred_probs_eoi_agg$made_cut_predicted_for_avg)), '\n')
```

```{r}
pred_probs_eoi_agg %>% filter(dg_id == 8825)
```

```{r}
test %>% filter(dg_id == 26096)
```
